\section{Methodology}
\label{sec:methodology}
\subsection{Counterfactual Paired Design}
Our evaluation uses \methodname, a within-item paired protocol. Each factual item is queried twice with identical answer options but different instruction framing: \control and \pressure. This design changes incentive context while holding item identity constant.

\para{Attribution rule.} Let $y_c$ and $y_p$ be binary correctness under control and pressure. On pressure-false items ($y_p=0$):
\begin{itemize}[leftmargin=*,itemsep=0pt,topsep=0pt]
    \item \epistemic if $y_c=0$ and $y_p=0$.
    \item \incentive-induced if $y_c=1$ and $y_p=0$.
\end{itemize}
No ambiguous cases occurred in this run.

\subsection{Data and Preprocessing}
We sample 120 items from \truthfulqa validation and 120 from \halueval, for 240 paired items total (seed 42). Data are loaded from local disk with `datasets.load\_from\_disk`.

For \truthfulqa, we use provided 
\mc choices and labels. For \halueval, we construct two-option \mc tasks (`right\_answer`, `hallucinated\_answer`) and randomize order. For each item, we mark one wrong option as the biased target in pressure prompts.

Data-quality checks report 0\% missingness in key fields, 817 \truthfulqa rows and 10,000 \halueval rows available, and \truthfulqa option counts ranging 2--13 (mean 5.04).

\subsection{Model, Prompts, and Inference}
We evaluate \gptmodel with `temperature=0.0`, `max\_tokens=220`, and `seed=42`. Prompt templates require a final answer line formatted as `FINAL: <LETTER>`. We parse that letter automatically and compute correctness per condition.

Runtime is about 8 minutes wall-clock. Hardware availability includes 2\,$\times$ RTX 3090 GPUs, though API evaluation runs on CPU-side orchestration.

\subsection{Metrics and Statistical Plan}
We report condition accuracy, false-rate risk difference (pressure minus control), and mechanism shares among pressure-false outputs. Paired comparisons use McNemar tests; multiple paired tests are corrected by Benjamini-Hochberg. Mechanism shares include 95\% bootstrap CIs (10,000 resamples).

\subsection{Baseline Detector and Transfer Evaluation}
We train \tfidfdet on pressure-response text to classify incentive-induced vs epistemic labels using TF-IDF 1--2 gram features and logistic regression (`max\_iter=2000`, balanced classes). We evaluate:
\begin{itemize}[leftmargin=*,itemsep=0pt,topsep=0pt]
    \item pooled random 70/30 split,
    \item train on \halueval and test on \truthfulqa,
    \item train on \truthfulqa and test on \halueval.
\end{itemize}
We report accuracy, precision, recall, F1, and \auroc.
